### 2023-12-24 sections/2-methods.tex

For a cluster $C$ with $|C|$ points, we begin by taking a random sample of $\sqrt{|C|}$ of its points, and computing pairwise distances between all points in this sample.
Using these distances, we compute the \emph{geometric median} of this sample; 
i.e., the point which minimizes the sum of distances to all other points in the sample.
This geometric median is the \emph{center} of $C$.

@@

Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-nearest neighbors search aims to find the set $S_q$ such that $|S_q| = k$ and $\forall x \in \textbf{X} \setminus S_q$, $\ f(x, q) > \max\{f(y, q): y \in S_q \}$;
i.e., for a given $k$, find the $k$ closest points to $q$ in $\textbf{X}$.
We also have the $\rho$-nearest neighbors search problem, which aims to find the set $\{x \in \textbf{X}: f(q, x) \leq \rho \}$;
that is, find all points in $\textbf{X}$ that are no more than a distance $\rho$ from $q$.

@@

The complexity bounds for Repeated $\rho$-NN rely on the assumption that the query point is sampled from the same distribution as the rest of the data, i.e. it arises from the same generative process as the rest of the data.
Given the uses of $k$-NN search in practice, this assumption is reasonable.
From this assumption, we can infer that the LFD at the query point does not differ significantly from the (harmonic) mean LFD of clusters near the query.

We find it useful to adopt the terminology used in \cite{ishaq2019clustered} and \cite{yu2015entropy}, and address \emph{tree-search} and \emph{leaf-search} separately.
Tree-search refers to the process of identifying clusters which have overlap with the query ball, i.e., which might contain one of the $k$ nearest neighbors. 
Leaf-search refers to the process of identifying the $k$ nearest neighbors among the points in those clusters identified by tree-search.

@@

We begin by letting $Q$ be a list of 3-tuples $(p, \delta^{+}_{p}, m)$, where $p$ is a cluster or a point, $\delta^{+}_{p}$ is the $\delta^{+}$ of $p$ as illustrated in Figure~\ref{fig:methods:deltas}, and $m$ is the multiplicity of $p$ in $Q$.
During the breadth-first traversal, for every cluster $C$ we encounter, we add $(C, \delta^{+}_{C}, |C| - 1)$ and $(c, \delta_{C}, 1)$ to $Q$, where $c$ is the center of $C$.

We then use the QuickSelect algorithm to find the element in $Q$ with the $k^{th}$ smallest $\delta^{+}$ while accounting for multiplicities, i.e. we find $\tau$, the smallest $\delta^{+}$ in $Q$ such that $\left| B_X(q, \tau) \right| \geq k$.

@@

We continue this process until $Q$ no longer contains any clusters.
We just use the QuickSelect algorithm once last time to reorder $Q$, to find the $k^{th}$ element, and to return the $k$ nearest neighbors.

@@

This algorithm is described in Algorithm~\ref{alg:methods:greedy-sieve}.
It terminates when $H$ is full and the top priority element in $H$ has $\delta$ less than the top priority element in $Q$, i.e. when the top priority element in $Q$ is farther from the query than the $k^{th}$ nearest neighbor in $H$.
This leaves $H$ containing exactly the $k$ nearest neighbors to the query.

@@

To elaborate on the augmentation process, we start with an original dataset from the ANN-benchmarks suite.
Let $X$ be the dataset, $d$ be its dimensionality, $\epsilon$ be a user-specified noise level and $m$ be a user-specified integer multiplier.
For each datum $x \in X$, we create $m$ new data points within a distance $\epsilon \cdot ||x||$ of $x$ where $||x||$ is the euclidean distance from $x$ to the origin.
We construct a random vector $r$ of $d$ dimensions in the hypersphere of radius $\epsilon$ centered at the origin.
We then add $r$ to $x$ to get a new point $x'$.
Since $||r|| = \epsilon$, we have that $||x - x'|| \leq \epsilon$, i.e. $x'$ is within a distance $\epsilon$ of $x$.
We repeat this process $m$ times for each $x \in X$ to get a new dataset $X'$.

>>>>

For a cluster $C$ with $|C|$ points, we begin by taking a random sample of $\sqrt{|C|}$ of its points, and computing pairwise distances between all points in this sample.
Using these distances, we compute the \emph{geometric median} of this sample; in other words, we find the point which minimizes the sum of distances to all other points in the sample.
This geometric median is the \emph{center} of $C$.

@@

Given a query $q$, along with a distance function $f$ defined on a dataset $\textbf{X}$, $k$-nearest neighbors search aims to find the set $S_q$ such that $|S_q| = k$ and $\forall x \in \textbf{X} \setminus S_q$, $\ f(x, q) > \max\{f(y, q): y \in S_q \}$;
in other words, for a given $k$, $k$-NN finds the $k$ closest points to $q$ in $\textbf{X}$.
We also have the $\rho$-nearest neighbors search problem, which aims to find the set $\{x \in \textbf{X}: f(q, x) \leq \rho \}$;
that is, find all points in $\textbf{X}$ that are no more than a distance $\rho$ from $q$.

@@

The complexity bounds for Repeated $\rho$-NN rely on the assumption that the query point is sampled from the same distribution as the rest of the data, or in other words, it arises from the same generative process as the rest of the data.
Given the uses of $k$-NN search in practice, this assumption is reasonable.
From this assumption, we can infer that the LFD at the query point does not differ significantly from the (harmonic) mean LFD of clusters near the query.

We find it useful to adopt the terminology used in \cite{ishaq2019clustered} and \cite{yu2015entropy}, and address \emph{tree-search} and \emph{leaf-search} separately.
Tree-search refers to the process of identifying clusters which have overlap with the query ball, or in other words, clusters which might contain one of the $k$ nearest neighbors. 
Leaf-search refers to the process of identifying the $k$ nearest neighbors among the points in those clusters identified by tree-search.

@@

We begin by letting $Q$ be a list of 3-tuples $(p, \delta^{+}_{p}, m)$, where $p$ is a cluster or a point, $\delta^{+}_{p}$ is the $\delta^{+}$ of $p$ as illustrated in Figure~\ref{fig:methods:deltas}, and $m$ is the multiplicity of $p$ in $Q$.
During the breadth-first traversal, for every cluster $C$ we encounter, we add $(C, \delta^{+}_{C}, |C| - 1)$ and $(c, \delta_{C}, 1)$ to $Q$, where $c$ is the center of $C$.
Recall that by the definitions of $\delta$ and $\delta^{+}$ given in Section~\ref{subsec:methods:the-search-problem}, $\delta_{C} = \delta_{c} = \delta^{+}_{c}$.

% TODO (Morgan): I don't like "kth smallest $\delta^{+}$ while accounting for multiplicities" 
We then use the QuickSelect algorithm to find the element in $Q$ with the $k^{th}$ smallest $\delta^{+}$ while accounting for multiplicities; in other words, we find $\tau$, the smallest $\delta^{+}$ in $Q$ such that $\left| B_X(q, \tau) \right| \geq k$.

@@

We continue this process until $Q$ no longer contains any clusters.
We then use the QuickSelect algorithm once last time to reorder $Q$, to find the $k^{th}$ element, and to return the $k$ nearest neighbors.

@@

This algorithm is described in Algorithm~\ref{alg:methods:greedy-sieve}.
It terminates when $H$ is full and the top priority element in $H$ has $\delta$ less than the top priority element in $Q$ (i.e., when the top priority element in $Q$ is farther from the query than the $k^{th}$ nearest neighbor in $H$).
This leaves $H$ containing exactly the $k$ nearest neighbors to the query.

@@

To elaborate on the augmentation process, we start with an original dataset from the ANN-benchmarks suite.
Let $X$ be the dataset, $d$ be its dimensionality, $\epsilon$ be a user-specified noise level and $m$ be a user-specified integer multiplier.
For each datum $x \in X$, we create $m$ new data points within a distance $\epsilon \cdot ||x||$ of $x$ where $||x||$ is the euclidean distance from $x$ to the origin.
We construct a random vector $r$ of $d$ dimensions in the hypersphere of radius $\epsilon$ centered at the origin.
We then add $r$ to $x$ to get a new point $x'$.
Since $||r|| = \epsilon$, we have that $||x - x'|| \leq \epsilon$ (i.e., $x'$ is within a distance $\epsilon$ of $x$).
We repeat this process $m$ times for each $x \in X$ to get a new dataset $X'$.
